Stack Safety is a Security Property

### 2021-04-12 tex/main.tex
Section 5 redone somewhat to take return targets into account

\section{Local State Encapsulation, Formally}
\label{sec:lse}

@@

\[\ann \in \ANNS ::= \callmap + \retmap\]
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \powerset(\ANNS)\]

\paragraph*{Call Depth and Domains}

At any given step of a program, some number of callers have pushed their frames
onto the stack, ``sealing'' that memory for later use; this number is the current
{\em call depth}. The call depth, as well as the association between stack addresses
and sealed frames, are dynamic and cannot be deduced from the state alone, so we will
track them over the course of an execution as additional context that informs the property.

Each component will be assigned to a {\em domain}, which may be {\em Outside} of the
stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed} by a particular
depth.

\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
  \times (\depth \in \mathbb{N}) \]

An initial context \(\context_0 = \domMap_0, 0\) where \(\domMap_0\) maps
the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
Then at each step we update the context according to the context-step function,
\(\stepsto_C\). We will define this in terms of the operations
\(\push{\domMap}{n}{\components}\), in which \(\components \subseteq \COMPONENTS\),
and \(\pop{\domMap}{n}\), pronounced ``push'' and ``pop'' and defined as follows.

\[
(\push{\domMap}{\depth}{\components}) ~ \component =
\begin{cases}
  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
                    \text{ and } \component \in \components \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]
\[
(\pop{\domMap}{\depth}) ~ \component =
\begin{cases}
  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth-1} \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]

Then we can define the context update relation. Let \(\sealcon\) be the sealing convention
that maps a machine state to a set of components, typically the set of addresses below
the stack pointer. The {\sc Default} rule applies only if neither of the others do.

\begin{center}
\begin{minipage}{.4\textwidth}
  \judgment{}
           {\(\mach,(\domMap,\depth),\{\callmap\} \stepstocon
             \push{\domMap}{\depth}{\sealcon ~ \mach},\depth+1\)}
\end{minipage}
\hspace*{0.1\textwidth}
\begin{minipage}{.4\textwidth}
  \judgment{}
           {\(\mach,(\domMap,\depth),\{\retmap\} \stepstocon
             \pop{\domMap}{\depth},\depth-1\)}
\end{minipage}
\end{center}

\begin{center}
  \judgment[Default]
           {}
           {\(\_,\context,\_ \stepstocon \context\)}
\end{center}

\subsection{Ultra-eager Integrity}

Our first property will be {\em ultra eager}, meaning that it will be quantified over each step
of execution from any initial state. The criterion for integrity is simple: sealed components do not
change.

\definition A system enjoys {\em ultra-eager integrity} if for any reachable state
\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
\pi_\mach(\mpcstatename') ~ \component\).

\subsection{Eager Integrity}

\definition Let \(\depth\) be any depth and \(f_\depth ~ \mach ~ (\_, \depth')\) hold when
\(\depth' \geq \depth\). A system enjoys {\em eager integrity} if each reachable segment
\(\MPCT \langle f_\depth \rangle\) satisfies \(\pi_m(\head(\MPCT)) ~ \component =
\pi_m(\last(\MPCT)) ~ \component\) for all \(\component\) where \(\domMap ~ \component =
\sealed{\_}\).

>>>>

\section{Stack Safety, Formally}
\label{sec:lse}

@@

\[\ann \in \ANNS ::= \callmap + \retmap\]
\[\codemap \in \CODEMAPS ::= \WORDS \rightharpoonup \FUNIDS \times \ANNS\]

When a function makes a call, it establishes that its frame should be
``sealed'' for later use. At any given step of a program, any number of
activations may have claimed memory in this way, but which cells are sealed
is dynamic information and cannot be deduced from the state alone. So it must
be tracked separately. We must also track the conditions under which a caller
considers itself to have been returned to.

Each system component will be assigned to a {\em domain}, which may be {\em Outside}
of the stack entirely, {\em Unsealed} and therefore accessible, or {\em Sealed}
at a particular depth. Each caller will register a {\em target} condition -- a
predicate on machine states that must hold for it to consider itself active.
A context is a pair of a ``domain map'' from components to domains, and a
``return target list,'' which is a stack of targets, one for each caller awaiting
return.

\[\stackDom \in \STKDOMS ::= \outside + \unsealed + \sealed{\depth}\]
\[\context \in \CONTEXTS ::= (\domMap \in \COMPONENTS \rightarrow \STKDOMS)
  \times (\rts \in \listT (\target \subseteq \MACHS)) \]

An initial context \(\context_0 = \domMap_0, []\) where \(\domMap_0\) maps
the stack to \(\unsealed\) and all other addresses and all registers to \(\outside\).
Then at each step we update the context according to the context-step function,
\(\stepsto_C\). We will define this in terms of the operations
\(\push{\domMap}{\depth}{\components}\), in which \(\depth \in \mathbb{Z}\) and
\(\components \subseteq \COMPONENTS\), and \(\pop{\domMap}{\depth}\), pronounced
``push'' and ``pop'' and defined as follows.

\[
(\push{\domMap}{\depth}{\components}) ~ \component =
\begin{cases}
  \sealed{\depth} & \text{if } \domMap ~ \component = \unsealed
                    \text{ and } \component \in \components \\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]
\[
(\pop{\domMap}{\depth}) ~ \component =
\begin{cases}
  \unsealed & \text{if } \domMap ~ \component = \sealed{\depth'} \text{ where }
              \depth' >= \depth\\
  \domMap ~ \component & \text{else} \\
\end{cases}
\]
%
Push seals a set of components with its depth argument, and pop unseals all components
that were as deep as its argument or deeper.

Now we can define the context update relation. Let \(\sealcon\) be the sealing convention
that maps a machine state to a set of components, typically the set of addresses below
the stack pointer. Similarly, \(\retcon ~ \mach ~ \mach'\) will be the return convention
that relates the call state \(\mach\) to return states \(\mach'\) that are valid returns.
The usual return convention is that \(\mach\) and \(\mach'\) agree on the stack pointer,
and \(\mach'\) has its program counter advanced by one instruction from \(\mach\).

The call rule pushes the components sealed by the sealing convention with the depth
of the current caller, and determines the return target by partially applying the
return convention to the machine state, and pushes that onto the targets stack.
The return rule checks whether the attempted return actually manages to reach a
return target, traversing down the stack until the first one it finds. Recall that
we are not discussing well-bracketed control flow, so the return might skip over
a waiting caller. When a target is reached, {\em all} sealed components at or above
its depth are unsealed by the pop operation. If no target is reached, nothing is
popped at all. These rules are given inductively, and apply in the listed order.

\judgmenttwo[Call]
            {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
            {\(\depth = |\rts|\)}
            {\(\mach,(\domMap,\rts) \stepsto_C
              \push{\domMap}{\depth}{\sealcon ~ \mach},(\retcon ~ \mach)::\rts\)}

\judgmentthree[ReturnFound]
              {\(\codemap ~ (\mach ~ \PCname) = \retmap\)}
              {\(\mach \stepsto \mach' \in \target\)}
              {\(\depth = |\rts|\)}
              {\(\mach,(\domMap,\target::\rts) \stepsto_C
                \pop{\domMap}{\depth},\rts\)}

\judgmentthree[ReturnRec]
              {\(\codemap ~ (\mach ~ \PCname) = \callmap\)}
              {\(\mach,(\domMap,\rts) \stepsto_C \_,\rts'\)}
              {\(\depth = |\rts'|\)}
              {\(\mach,(\domMap,\_::\rts) \stepsto_C
                \pop{\domMap}{\depth},\rts'\)}

\judgment[Default]
         {}
         {\(\_,\context \stepsto_C \context\)}

When we build machine traces with this step function, we always know how deep in the
call stack we are and which components are sealed. We may now proceed to state our properties.
Our first property will be {\em stepwise stack integrity}, stepwise meaning that it will be
quantified over each step of execution from any initial state. The criterion for integrity
is simple: sealed components do not change.

\definition A system enjoys {\em stepwise stack integrity} if for any reachable state
\(\mpcstatename\), \(\mpcstatename'\) such that \(\mpcstatename \stepstocon \mpcstatename'\),
and component \(\component\), if \(\pi_\context(\mpcstatename) = \domMap,\_\) and
\(\domMap ~ \component = \sealed{\_}\), then \(\pi_\mach(\mpcstatename) ~ \component =
\pi_\mach(\mpcstatename') ~ \component\).

Next we consider {\em lockstep stack confidentiality}. Here it behooves us to factor out
the general notion of lockstep confidentiality independent of a particular context model.
